---
title: "anova"
output: html_document
date: "2022-11"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(ggtext)
```


## ANOVA 


**One-way ANOVA:**

$k =$ number of populations or treatments being compared

$\mu_1 =$ mean of population 1 or true average response when treatment 1 is applied.

...

$\mu_k =$ mean of population $k$ or true average response when treatment $k$ is applied.
  
**Assumptions:**


- For each population, response variable is normally distributed
- Variance of response variable, $\sigma^2$ is the same for all the populations
- The observations must be independent


**Hypotheses:**

$$H_o: \mu_1 = \mu_2 = ... = \mu_k$$

$$H_a: \mu_i \neq \mu_j$ for $i \neq j$$

**Notation:**

$y_{ij}$ is the $j^{th}$ observed value from the $i^{th}$ population/treatment.

Total mean:  $\overline{y}_{i\cdot} = \frac{y_{i\cdot}}{n_i} = \frac{\sum_{j=1}^{n_i}y_{ij}}{n_i}$
      
Total sample size:  $n = n_1 + n_2 + ... + n_k$ 
  	
Grand total:  $y_{\cdot\cdot} = \sum_{i=1}^{k}\sum_{j=1}^{n_i}y_{ij}$
  	
Grand mean:  $\overline{y}_{\cdot\cdot} = \frac{y_{\cdot\cdot}}{n} = \frac{\sum_{i=1}^{k}\sum_{j=1}^{n_i}y_{ij}}{n}$


$s^2 = \frac{\sum_{j=1}^{k}(n_i-1)s_i^2}{n-k}$ = MSE, where	$s_i^2 = \frac{\sum_{j=1}^{n_i}(y_{ij}-\overline{y}_{i\cdot})^2}{n_i-1}$\\

**The ANOVA Table:**


|     Source of Variation | df | Sum of Squares | Mean Square | F-ratio |
|:---|:---:|:---:|:---:|:---:|:---:|
|     Treatment | $k-1$ | SSTr | MSTr = $\frac{\textrm{SSTr}}{k-1}$ | $\frac{\textrm{MSTr}}{\textrm{MSE}}$ |
|     Error | $n-k$ | SSE | MSE = $\frac{\textrm{SSE}}{n-k}$ |  |
|     Total | $n-1$ | SST |  |  |

SST = SSTr + SSE
      
SST = $\sum_{i=1}^{k}\sum_{j=1}^{n_i}(y_{ij}-\overline{y}_{\cdot\cdot})^2 = \sum_{i=1}^{k}\sum_{j=1}^{n_i}y_{ij}^2-\frac{1}{n}y_{\cdot\cdot}^2$

SSTr = $\sum_{i=1}^{k}\sum_{j=1}^{n_i}(\overline{y}_{i\cdot}-\overline{y}_{\cdot\cdot})^2 = \sum_{i=1}^{k}\frac{1}{n_i}y_{i\cdot}^2-\frac{1}{n}y_{\cdot\cdot}^2$

SSE = $\sum_{i=1}^{k}\sum_{j=1}^{n_i}(y_{ij}-\overline{y}_{i\cdot})^2= \sum_{i=1}^{k}\sum_{j=1}^{n_i}y_{ij}^2-\sum_{i=1}^{k}\frac{y_{i\cdot}^2}{n_i} = \sum_{i=1}^{k}(n_i-1)s_i^2$

**Test Statistic:**

$F_{obs} = \frac{\textrm{MSTr}}{\textrm{MSE}} \sim F_{v_1, v_2} \qquad  v1 = df(\textrm{SSTr}) = k-1 \qquad v2 = df(\textrm{SSE}) = n-k$
	
Reject $H_o$ if $F_{obs} \geq F_{\alpha, v_1, v_2}$

```{r anova, echo=FALSE}
inputPanel(
  sliderInput("varianza", label = "Varianza:",
              min = 0.2, max = 6, value = 3, step = 0.2),
  sliderInput("nivel_1", label = "Nivel medio del tratamiento 1:",
              min = 0, max = 5, value = 0.0, step = 0.2), 
  sliderInput("nivel_2", label = "Nivel medio del tratamiento 2:",
              min = 0, max = 5, value = 0.2, step = 0.2),
  sliderInput("nivel_3", label = "Nivel medio del tratamiento 3:",
              min = 0, max = 5, value = 0.4, step = 0.2),
  sliderInput("n", label = "n:",
              min = 10, max = 30, value = 15, step = 1), 
  checkboxInput(inputId="Intervalo",
                           label="Intervalo de Confianza", 
                           value = F), 
  checkboxInput(inputId="Media",
                           label="Promedio Global", 
                           value = F)
)

renderPlot({

  set.seed(128498)
  n <- input$n
  y <- rnorm(3*n)
  tratamiento <- rep(1:3,each = n)
  Datos <- data.frame(tratamiento=as.factor(tratamiento),y)  %>% 
    group_by(tratamiento) %>% 
    mutate(y = (y- mean(y))/sd(y)) %>% 
    ungroup()
  Datos$y <- Datos$y*sqrt(input$varianza) + 
    rep(c(input$nivel_1,input$nivel_2,input$nivel_3),each = n)
  resumen <- Datos %>% 
    Rmisc::summarySE(data =., measurevar="y", groupvars=c("tratamiento"))
  base2 <- Datos %>%
    ggplot(aes(tratamiento,y)) +
    geom_errorbar(data=resumen,aes(ymin=y, ymax=y), width=.3, color ="blue") +
    ylim(-5, 10) + 
    theme_bw()
  
  if (input$Intervalo) 
{    g1 <- base2 + geom_errorbar(data=resumen,
                                aes(ymin=y-ci, ymax=y+ci), 
                                width=.1, 
                                color ="blue")}
  else
{    g1 <- base2 + geom_dotplot(binaxis='y', 
                               stackdir='center', 
                               dotsize=.3, 
                               colour = tratamiento, 
                               fill = tratamiento)}
 
    if (input$Media) 
    g1 <- g1 + geom_hline(yintercept = mean(Datos$y, na.rm=TRUE))

  base <-   ggplot() +
  xlim(-5, 10)
  
  g2 <- base + 
    geom_function(aes(colour = "Tratamiento 1"),
                  fun = dnorm, 
                  args = list(mean = input$nivel_1, sd =sqrt(input$varianza)),
                  color = "black")+
    geom_function(aes(colour = "Tratamiento 2"),
                  fun = dnorm, 
                  args = list(mean = input$nivel_2, sd =sqrt(input$varianza)),
                  color = "red")+    
    geom_function(aes(colour = "Tratamiento 3"),
                  fun = dnorm, 
                  args = list(mean = input$nivel_3, sd =sqrt(input$varianza)),
                  color = "green")+    
    coord_flip() + 
    theme_bw()
  

  tabla <- Datos %>% lm(data = ., y ~ tratamiento) %>% anova() %>% round(2)
  
  text1 <- gridExtra::tableGrob(replace(tabla, is.na(tabla), "")) 
    
  (g2+ g1 + text1)+ plot_layout(widths = c(1, 3, 2))
})
```

