---
title: "anova"
output: html_document
date: "2022-11"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(ggtext)
```


## ANOVA 
$a$


**One-way ANOVA:**

$k =$ number of populations or treatments being compared

$\mu_1 =$ mean of population 1 or true average response when treatment 1 is applied.

...

$\mu_k =$ mean of population $k$ or true average response when treatment $k$ is applied.
  
**Assumptions:**


- For each population, response variable is normally distributed
- Variance of response variable, $\sigma^2$ is the same for all the populations
- The observations must be independent


**Hypotheses:**

$H_o: \mu_1 = \mu_2 = ... = \mu_k$

$H_a: \mu_i \neq \mu_j$ for $i \neq j$

**Notation:**

$y_{ij}$ is the $j^{th}$ observed value from the $i^{th}$ population/treatment.

Total mean:  $\overline{y}_{i\cdot} = \frac{y_{i\cdot}}{n_i} = \frac{\sum_{j=1}^{n_i}y_{ij}}{n_i}$
      
Total sample size:  $n = n_1 + n_2 + ... + n_k$ 
  	
Grand total:  $y_{\cdot\cdot} = \sum_{i=1}^{k}\sum_{j=1}^{n_i}y_{ij}$
  	
Grand mean:  $\overline{y}_{\cdot\cdot} = \frac{y_{\cdot\cdot}}{n} = \frac{\sum_{i=1}^{k}\sum_{j=1}^{n_i}y_{ij}}{n}$


$s^2 = \frac{\sum_{j=1}^{k}(n_i-1)s_i^2}{n-k}$ = MSE, where	$s_i^2 = \frac{\sum_{j=1}^{n_i}(y_{ij}-\overline{y}_{i\cdot})^2}{n_i-1}$\\

**The ANOVA Table:**


|     Source of Variation | df | Sum of Squares | Mean Square | F-ratio |
|:---|:---:|:---:|:---:|:---:|:---:|
|     Treatment | $k-1$ | SSTr | MSTr = $\frac{\textrm{SSTr}}{k-1}$ | $\frac{\textrm{MSTr}}{\textrm{MSE}}$ |
|     Error | $n-k$ | SSE | MSE = $\frac{\textrm{SSE}}{n-k}$ |  |
|     Total | $n-1$ | SST |  |  |



SST = SSTr + SSE
      
SST = $\sum_{i=1}^{k}\sum_{j=1}^{n_i}(y_{ij}-\overline{y}_{\cdot\cdot})^2 = \sum_{i=1}^{k}\sum_{j=1}^{n_i}y_{ij}^2-\frac{1}{n}y_{\cdot\cdot}^2$

SSTr = $\sum_{i=1}^{k}\sum_{j=1}^{n_i}(\overline{y}_{i\cdot}-\overline{y}_{\cdot\cdot})^2 = \sum_{i=1}^{k}\frac{1}{n_i}y_{i\cdot}^2-\frac{1}{n}y_{\cdot\cdot}^2$

SSE = $\sum_{i=1}^{k}\sum_{j=1}^{n_i}(y_{ij}-\overline{y}_{i\cdot})^2= \sum_{i=1}^{k}\sum_{j=1}^{n_i}y_{ij}^2-\sum_{i=1}^{k}\frac{y_{i\cdot}^2}{n_i} = \sum_{i=1}^{k}(n_i-1)s_i^2$


  
**Test Statistic:**

$F_{obs} = \frac{\textrm{MSTr}}{\textrm{MSE}} \sim F_{v_1, v_2}$   $v1 = df(\textrm{SSTr}) = k-1$
	
$v2 = df(\textrm{SSE}) = n-k$
	
Reject $H_o$ if $F_{obs} \geq F_{\alpha, v_1, v_2}$

  

```{r anova, echo=FALSE}
inputPanel(
  sliderInput("varianza", label = "Varianza:",
              min = 0.2, max = 4, value = 1, step = 0.2),
   sliderInput("nivel_1", label = "Nivel medio del tratamiento 1:",
              min = 0.5, max = 5, value = 1, step = 0.5), 
   sliderInput("nivel_2", label = "Nivel medio del tratamiento 2:",
              min = 0.5, max = 5, value = 1.5, step = 0.5),
   sliderInput("nivel_3", label = "Nivel medio del tratamiento 3:",
              min = 0.5, max = 5, value = 2, step = 0.5)  
)

renderPlot({

  set.seed(128498)
  n <- 15
  error <- rnorm(3*n,0,sqrt(input$varianza))
  tratamiento <- rep(1:3,each = n)
  y <- rep(c(input$nivel_1,input$nivel_2,input$nivel_3),each = n) + error
  Datos <- data.frame(tratamiento=as.factor(tratamiento),y) 
  g1 <- Datos %>%
    ggplot(aes(tratamiento,y))+
    geom_boxplot() +
    geom_dotplot(binaxis='y', stackdir='center', dotsize=1,colour = tratamiento) +
    ylim(-5, 10) + theme_bw()
  base <-   ggplot() +
  xlim(-5, 10)
  g2 <- base + 
    geom_function(aes(colour = "Tratamiento 1"),
                  fun = dnorm, 
                  args = list(mean = input$nivel_1, sd =sqrt(input$varianza)),
                  color = "black")+
    geom_function(aes(colour = "Tratamiento 2"),
                  fun = dnorm, 
                  args = list(mean = input$nivel_2, sd =sqrt(input$varianza)),
                  color = "red")+    
    geom_function(aes(colour = "Tratamiento 3"),
                  fun = dnorm, 
                  args = list(mean = input$nivel_3, sd =sqrt(input$varianza)),
                  color = "green")+    
    coord_flip() + theme_bw()
  
  string <- "Veamos ahora que se puede hacer"
  text1 <- ggplot(data = tibble(x = 0, y = 1, label = string)) +
  aes(x = x, y = y, label = label) +
  geom_textbox(
    box.color = NA,
    fill = NA,
    width = unit(10, "cm"),
    hjust = 0,
    vjust = 1
  ) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme_void() +
  theme(
    plot.margin = margin(0, 0, 0, 0)
  )
  tabla <- Datos %>% lm(data = ., y ~ tratamiento) %>% anova() %>% round(2)
  
  text1 <- gridExtra::tableGrob(tabla) 
    
  (g2+ g1 + text1)+ plot_layout(widths = c(1, 3, 2))
})
```

